---
title: "Results"
author: '180000622'
date: "2023-08-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(foreach)
library(EMD)
setwd("~/MSc_Thesis_Supplementary_Material/functions")
source('significance_test_functions.R')
setwd("~/MSc_Thesis_Supplementary_Material")
```

```{r}
# source prepared data objects
long_left_IMFs <- read_rds('long_series_IMFs.rds')
scaled_long_left_IMFs <- read_rds('scaled_long_series_IMFs.rds')

scaled_price_IMFs <- read_rds('scaled_price_IMFs.rds')
unscaled_price_IMFs <- read_rds('unscaled_price_IMFs.rds')
```

First, let's consider what files we have:

```{r}
tickers <- names(long_left_IMFs)

n_forecasts <- lapply(tickers, function(X) {
  forecasts <- long_left_IMFs[[X]]$IMF_data
  
  lengths <- foreach(forecast = forecasts) %do% {
    forecast$generation_dates %>% length()
  }
  
  mean_len <- mean(unlist(lengths, use.names = FALSE))
  range_len <- range(unlist(lengths, use.names = FALSE))
  
  return(data.frame(ticker = X,
                    num_forecasts = length(forecasts),
                    mean_length = mean_len,
                    low_range = range_len[1],
                    up_range = range_len[2]))
})

n_forecasts_df <- bind_rows(n_forecasts)
```
```{r}
number_of_series <- n_forecasts_df$num_forecasts %>% sum()
number_of_series
```
# Wu signif tests:

First get energy density profiles for each expectations series
```{r}
# select the number of series that are IMFs
num_IMFs <- log2(200) %>% round() - 1

en_dens_score <- foreach(ticker = tickers) %do% {
  IMF_df <- scaled_long_left_IMFs[[ticker]]$IMF_data
  
  n_merged_sets <- length(IMF_df)
  
  en_dens_period_generation <- vector('list', )
  for (i in 1:n_merged_sets) {
    IMFs <- IMF_df[[i]]$IMFs
    en_dens_period_generation[[i]] <- foreach(series = IMFs, .final = function(x) {bind_rows(x)}) %do% {
                                      en_den_period_df <- data.frame(IMFs = colnames(series[,1:num_IMFs]),
                                      energy_density = energy_density_func(series[,1:num_IMFs]) %>% 
                                       unlist(use.names = FALSE),
                                      mean_periods = lapply(series[,1:num_IMFs], mean_period_func) %>% unlist(use.names = FALSE) )}
  }
  en_dens_period_generation
}
en_dens_score <- set_names(en_dens_score, tickers)
```

Significane Test 1: g+1 for all series

To get the g+1 energy density profiles for each series we can extract the first 7 observations from each energy density profile

```{r}
# this set of loops selects the first 7 observations from each energy density profile
# for each company and binds them all together to form an energy density profile for every 
# g+1 observation in our long series subset fo the expectation data
g1_en_dens <- foreach(ticker = tickers, .combine = bind_rows) %do% {
  ticker_en_den <- en_dens_score[[ticker]]
  
  g1 <- foreach(profile = ticker_en_den, .combine = bind_rows) %do% {
    # select first 7 rows of each profile
    profile[1:7, ]
  }
}

```

## g+1 Wu test: 
```{r}

  # set variables for confidence intervals
  # We will set N to be the mean length of the observed forecasts

  N <- mean(n_forecasts_df$mean_length) %>% round()
  x_series <- seq(0, max(log(g1_en_dens$mean_periods)), 0.1)

  CI1 <- data.frame(x = x_series,
                 CI = confidence_intervals_func(x_series, alpha = 0.01, N = N))

  CI5 <- data.frame(x = x_series,
                 CI = confidence_intervals_func(x_series, alpha = 0.05, N = N))


  # plot
  g1_en_dens %>% ggplot() +
    # plot scaled expectations
    geom_point(aes(x = log(mean_periods), 
                 y = log(energy_density),
                 color = IMFs), size = 1.5, shape = 'triangle') +

  geom_line(data = CI1, aes(x = x, y = CI.lower), linetype = 'dashed') +
  geom_line(data = CI1, aes(x = x, y = CI.upper), linetype = 'dashed') +
  
  # add 5% confidence interval
   geom_line(data = CI5, aes(x = x, y = CI.lower), linetype = 'dashed', color = 'red') +
  geom_line(data = CI5, aes(x = x, y = CI.upper), linetype = 'dashed', color = 'red') +
    
    labs(x = expression(ln(bar(T))),
         y = expression(ln(E)),
         color = NULL) +
  theme_minimal()
```

# all g+h Wu test

Now let's plot a Wu test for all the g+h series
```{r}
# this set of loops selects all observations from each energy density profile
# for each company and binds them all together to form an energy density profile for every 
# g+h observation in our long series subset fo the expectation data
gh_en_dens <- foreach(ticker = tickers, .combine = bind_rows) %do% {
  ticker_en_den <- en_dens_score[[ticker]]
  
  gh <- foreach(profile = ticker_en_den, .combine = bind_rows) %do% {
    # select first 7 rows of each profile
    profile
  }
}
```

```{r}

  # set variables for confidence intervals
  # We will set N to be the mean length of the observed forecasts

  N <- mean(n_forecasts_df$mean_length) %>% round()
  x_series <- seq(0, max(log(gh_en_dens$mean_periods)), 0.1)

  CI1 <- data.frame(x = x_series,
                 CI = confidence_intervals_func(x_series, alpha = 0.01, N = N))

  CI5 <- data.frame(x = x_series,
                 CI = confidence_intervals_func(x_series, alpha = 0.05, N = N))


  # plot
  gh_en_dens %>% ggplot() +
    # plot scaled expectations
    geom_point(aes(x = log(mean_periods), 
                 y = log(energy_density),
                 color = IMFs), size = 1.5, shape = 'triangle',
               alpha = 0.5) +

  geom_line(data = CI1, aes(x = x, y = CI.lower), linetype = 'dashed') +
  geom_line(data = CI1, aes(x = x, y = CI.upper), linetype = 'dashed') +
  
  # add 5% confidence interval
   geom_line(data = CI5, aes(x = x, y = CI.lower), linetype = 'dashed', color = 'red') +
  geom_line(data = CI5, aes(x = x, y = CI.upper), linetype = 'dashed', color = 'red') +
    
    labs(x = expression(ln(bar(T))),
         y = expression(ln(E)),
         color = NULL) +
  theme_minimal()
```
# Stock Price Wu Test

We need to get the energy density profile of the stock price IMFs

```{r}
en_dens_price <- foreach(ticker = tickers, .combine = bind_rows) %do% {
  IMF_df <- scaled_price_IMFs[[ticker]]$price_IMFs
  
  n_merged_sets <- length(IMF_df)
  
  en_dens_period_generation <- vector('list', )

  en_dens_period_generation <- foreach(series = IMF_df, .combine = bind_rows) %do% {
                                  en_den_period_df <- data.frame(IMFs = colnames(series[,1:num_IMFs]),
                                  energy_density = energy_density_func(series[,1:num_IMFs]) %>% 
                                  unlist(use.names = FALSE),
                                  mean_periods = lapply(series[,1:num_IMFs], mean_period_func) %>% unlist(use.names = FALSE))}
  en_dens_period_generation
}
```

```{r}

  # set variables for confidence intervals
  # We will set N to be the mean length of the observed forecasts

  N <- mean(n_forecasts_df$mean_length) %>% round()
  x_series <- seq(0, max(log(en_dens_price$mean_periods)), 0.1)

  CI1 <- data.frame(x = x_series,
                 CI = confidence_intervals_func(x_series, alpha = 0.01, N = N))

  CI5 <- data.frame(x = x_series,
                 CI = confidence_intervals_func(x_series, alpha = 0.05, N = N))


  # plot
  en_dens_price %>% ggplot() +
    # plot scaled expectations
    geom_point(aes(x = log(mean_periods), 
                 y = log(energy_density),
                 color = IMFs), size = 1.5, shape = 'square',
               alpha = 0.75) +

  geom_line(data = CI1, aes(x = x, y = CI.lower), linetype = 'dashed') +
  geom_line(data = CI1, aes(x = x, y = CI.upper), linetype = 'dashed') +
  
  # add 5% confidence interval
   geom_line(data = CI5, aes(x = x, y = CI.lower), linetype = 'dashed', color = 'red') +
  geom_line(data = CI5, aes(x = x, y = CI.upper), linetype = 'dashed', color = 'red') +
    
    labs(x = expression(ln(bar(T))),
         y = expression(ln(E)),
         color = NULL) +
  theme_minimal()
```

# Wu test for Price Change:
```{r}
en_dens_change <- foreach(ticker = tickers, .combine = bind_rows) %do% {
  IMF_df <- scaled_price_IMFs[[ticker]]$change_IMFs
  
  n_merged_sets <- length(IMF_df)
  
  en_dens_period_generation <- vector('list', )

  en_dens_period_generation <- foreach(series = IMF_df, .combine = bind_rows) %do% {
                                  en_den_period_df <- data.frame(IMFs = colnames(series[,1:num_IMFs]),
                                  energy_density = energy_density_func(series[,1:num_IMFs]) %>% 
                                  unlist(use.names = FALSE),
                                  mean_periods = lapply(series[,1:num_IMFs], mean_period_func) %>% unlist(use.names = FALSE))}
  en_dens_period_generation
}
```

```{r}

  # set variables for confidence intervals
  # We will set N to be the mean length of the observed forecasts

  N <- mean(n_forecasts_df$mean_length) %>% round()
  x_series <- seq(0, max(log(en_dens_change$mean_periods)), 0.1)

  CI1 <- data.frame(x = x_series,
                 CI = confidence_intervals_func(x_series, alpha = 0.01, N = N))

  CI5 <- data.frame(x = x_series,
                 CI = confidence_intervals_func(x_series, alpha = 0.05, N = N))


  # plot
  en_dens_change %>% ggplot() +
    # plot scaled expectations
    geom_point(aes(x = log(mean_periods), 
                 y = log(energy_density),
                 color = IMFs), size = 1.5,
               alpha = 0.75) +

  geom_line(data = CI1, aes(x = x, y = CI.lower), linetype = 'dashed') +
  geom_line(data = CI1, aes(x = x, y = CI.upper), linetype = 'dashed') +
  
  # add 5% confidence interval
   geom_line(data = CI5, aes(x = x, y = CI.lower), linetype = 'dashed', color = 'red') +
  geom_line(data = CI5, aes(x = x, y = CI.upper), linetype = 'dashed', color = 'red') +
    
    labs(x = expression(ln(bar(T))),
         y = expression(ln(E)),
         color = NULL) +
  theme_minimal()
```

# Frequency/ Frequency plots

To construct the frequency/frequency plot we need to do the hilbert transform to get 
Instantaneous frequencies and amplitudes

## Price Hilbert Transform
```{r}
# loop over each ticker and do the hilbert transform on each set of IMFs

amp_freq_price <- foreach(ticker = tickers) %do% {
  ticker_price_IMFs <- unscaled_price_IMFs[[ticker]]$price_IMFs
  
  # hilbertspec does the hilbert transform and computes the instantaneous frequency
  # and amplitude for each IMF
  amp_freq_list <- foreach(gen_period = ticker_price_IMFs) %do% {
    EMD::hilbertspec(gen_period[,1:num_IMFs]) 
  }
}
amp_freq_price <- set_names(amp_freq_price, tickers)
```
## Change Hilbert Transform
```{r}
# loop over each ticker and do the hilbert transform on each set of IMFs

amp_freq_change <- foreach(ticker = tickers) %do% {
  ticker_change_IMFs <- unscaled_price_IMFs[[ticker]]$change_IMFs
  
  # hilbertspec does the hilbert transform and computes the instantaneous frequency
  # and amplitude for each IMF
  amp_freq_list <- foreach(gen_period = ticker_change_IMFs) %do% {
    EMD::hilbertspec(gen_period[,1:num_IMFs]) 
  }
}
amp_freq_change <- set_names(amp_freq_change, tickers)
```

## g+h Hilbert Transform
```{r}
amp_freq_score <- foreach(ticker = tickers) %do% {
  ticker_score_IMFs <- long_left_IMFs[[ticker]]$IMF_data
  
  # loop over the different merged foreacasts
  amp_freq_list <- foreach(gen_period = ticker_score_IMFs) %do% {
    # loop over the different g+h series within each merged forecast
    IMFs <- gen_period$IMFs
    amp_freq_sub_list <- foreach(series = IMFs) %do% {
      amps_freqs <- EMD::hilbertspec(series[ ,1:num_IMFs])
    }
  }
}
amp_freq_score <- set_names(amp_freq_score, tickers)
```

## g+1 frequencies

To compare all the g+1 frequencies against their respective price and price changes we need to make a data frame that pairs each series against each other for their respective observations.

```{r}
# this chunk selects all the g1 instantaneous frequencies and places them in a data frame with
# the corresponding price and change frequencies

# IMF colnames for IMF column in output df
IMF_names <- names(unscaled_price_IMFs$ADM[[1]][[1]])[1:num_IMFs]

g1_frequency <- foreach(ticker = tickers) %do% {
  
  af_price_ticker <- amp_freq_price[[ticker]]
  af_change_ticker <- amp_freq_change[[ticker]]
  af_score_ticker <- amp_freq_score[[ticker]]
  
  instant_freq <- foreach(i = seq(1:length(af_score_ticker))) %do% {
    # make data frame by extracting IF matrix and stacking columns to form one long vector. Add
    # an IMF column to identify the vectors. We must transpose the IF matrices so that the as.vector
    # function stacks the rows keeping the ordering of IMF names the same
    IF <- data.frame(IMF = rep(IMF_names, num_IMFs * nrow(af_score_ticker[[i]][[1]]$instantfreq)), 
                     g1 = af_score_ticker[[i]][[1]]$instantfreq %>% t() %>% as.vector(),
                     price = af_price_ticker[[i]]$instantfreq %>% t() %>% as.vector(),
                     change = af_change_ticker[[i]]$instantfreq %>% t() %>% as.vector())
  }
}

```

# Frequency Scatterplots
```{r}
g1_bound <- g1_frequency %>% bind_rows()

means_freq <- g1_bound %>% group_by(IMF) %>% summarise(score_mean = mean(`g1`),
                                                         price_mean = mean(price),
                                                         change_mean = mean(change))

```
```{r}
g1_price <- g1_bound %>% ggplot(aes(x = `g1`, y = price, color = IMF)) +
  geom_point() + geom_abline(intercept = 0, slope = 1, linetype = 'dashed') +
  geom_point(data = means_freq, aes(x = score_mean, y = price_mean, color = IMF), color = 'black', shape = "X", size = 3) +
  theme_minimal() +
  labs(y = 'Price Close',
       x = 'g+1') +
  theme(legend.position = 'none')
g1_price

g1_change <- g1_bound %>% ggplot(aes(x = `g1`, y = change, color = IMF)) +
  geom_point() + geom_abline(intercept = 0, slope = 1, linetype = 'dashed') +
  geom_point(data = means_freq, aes(x = score_mean, y = change_mean, color = IMF), color = 'black', shape = "X", size = 3) +
  theme_minimal() +
  labs(y = 'Price Change',
       x = 'g+1') +
  theme(legend.position = 'none')

g1_change

price_change <- g1_bound %>% ggplot(aes(x = price, y = change, color = IMF)) +
  geom_point() + geom_abline(intercept = 0, slope = 1, linetype = 'dashed') +
  geom_point(data = means_freq, aes(x = price_mean, y = change_mean, color = IMF), color = 'black', shape = "X", size = 3) +
  theme_minimal() +
  labs(y = 'Price Change',
       x = 'Price Close')
price_change
```

```{r}
g1_bound %>% ggplot() + geom_histogram(aes(g1, fill = IMF), bins =100) + facet_wrap(~IMF, scales = 'free')
g1_bound %>% ggplot() + geom_histogram(aes(price, fill = IMF), bins = 100) + facet_wrap(~IMF, scales = 'free')
g1_bound %>% ggplot() + geom_histogram(aes(change,  fill = IMF), bins = 100) + facet_wrap(~IMF, scales = 'free')
```
```{r}
cor.test(g1_bound$g1, g1_bound$price)
cor.test(g1_bound$g1, g1_bound$change)
cor.test(g1_bound$price, g1_bound$change)
```

```{r}
cor.test(g1_bound$g1[g1_bound$IMF == 'IMF 1'], g1_bound$price[g1_bound$IMF == 'IMF 1'])
cor.test(g1_bound$g1[g1_bound$IMF == 'IMF 2'], g1_bound$price[g1_bound$IMF == 'IMF 2'])
cor.test(g1_bound$g1[g1_bound$IMF == 'IMF 3'], g1_bound$price[g1_bound$IMF == 'IMF 3'])
cor.test(g1_bound$g1[g1_bound$IMF == 'IMF 4'], g1_bound$price[g1_bound$IMF == 'IMF 4'])
cor.test(g1_bound$g1[g1_bound$IMF == 'IMF 5'], g1_bound$price[g1_bound$IMF == 'IMF 5'])
cor.test(g1_bound$g1[g1_bound$IMF == 'IMF 6'], g1_bound$price[g1_bound$IMF == 'IMF 6'])
cor.test(g1_bound$g1[g1_bound$IMF == 'IMF 7'], g1_bound$price[g1_bound$IMF == 'IMF 7'])

```
